---
title: "Introduction to the FRAS package"
author: "Dinh Dong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the FRAS package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, include = FALSE}
library(FRAS)
library(dplyr)
library(maps)
```

This vignette gives an introduction to the FRAS package. The package contains a set of functions used for analysis of data from FARS - Fatality Analysis Reporting System. FARS is a nationwide census providing NHTSA, Congress and the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. We can get from:  [[1](http://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS))]

## Package functions

There are three exported functions available to users:

- `fars_read()`
- `fars_summarize_years()`
- `fars_map_state()`

We will use data from FARS, years 2013-2015, that can be found in the package directory under `\dataset` folder for create the examples.

## Read FARS data files

The function `fars_read()` provides a way how to read data from a file path provided as an argument. It returns a `tibble` with loaded data:

```{r fars_read_example}
filename <- system.file("extdata/accident_2014.csv.bz2", package = "FRAS")
fars_read(filename)
```

## Summarize the number of accidents

The 2nd function `fars_summarize_years()` input a list of years with integer values as an argument. If the input files with data are available, it returns a data frame with number of accidents per year and month. The data files need to be located in the working directory.

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "FRAS"))
fars_summarize_years(2013:2015)
```

## Plot accident locations

The last function is `fars_map_state()` needs a state number and year as arguments. It returns a plot of the state with accident locations associated with years. This is an example showing how to use this function.

```{r fars_map_state_example}
setwd(system.file("extdata", package = "FRAS"))
fars_map_state(10, 2013)
```
